---
title: "Assignment01"
author: "Jessica Tang"
date: "9/10/2020"
output:
  html_document:
   toc: true
   toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

## Housing and Food Security in NY State, by Household Type

I loaded the following libraries:

```{r libraries}
library(tidycensus)
library(tidyverse)
library(dplyr)
```

(P.S. I used "setup" code from the written tutorial to hide the warnings and messages I get when knitting my markdown file.)

Then, I used `data("pums_variables")` to see a dataframe of variables available in PUMS.

I would like to see the proportion and types of households in NY State that are burdened and are on food assistance, to answer the following questions:

1. What is the prevalence of this double-barrier situation (i.e., housing cost-burdened and on food assistance)?
2. Are renter-occupied households more or less likely to be on food assistance than owner-occupied households are, regardless of whether these homeowners own their homes "free and clear" or whether they have a mortgage?
3. How does the prevalence of this double-barrier situation vary by the household's English-speaking ability?
4. What is the relationship between rent or owner housing costs, household income, and number of persons in the household? At what rate would income have to increase in order for the household to not be burdened by housing costs, and how does this change as the number of persons in household increases or decreases?

For each renter-occupied and owner-occupied household types, the 4 main situations would be:

* not burdened (GRPIP or OCPIP is < 30%) & not on food assistance
* not burdened & on food assistance
* burdened (GRPIP or OCPIP is >= 30%) & not on food assistance
* burdened & on food assistance

with additional variations based on household income, English-speaking ability, and number of persons in household.

**Sample population:** Households in NY State that are renter-occupied or owner-occupied (either "free and clear" or with a mortgage; vacant households are excluded). I excluded households that are "occupied without payment of rent" (e.g., caretakers, superintendents), because I imagine this is a smaller subset of the NY State population as a whole. Plus, housing cost is one of the very variables I would like to examine.

### For Renters

Variables:

1. Tenure (TEN), categorical
2. Yearly food stamp/Supplemental Nutrition Assistance Program (SNAP) recipiency (FS), categorical
3. Limited English-speaking household (LNGI), categorical
4. Gross rent as a percentage of household income past 12 months (GRPIP), categorical in the sense that < 30% is non-burdened and >= 30% is burdened
5. Gross rent (GRNTP), continuous
6. Household income (HINCP), continuous
7. Number of persons in this household (NP), continuous

Number of observations: 60,512

```{r hh_renter_data, results = 'hide'}
hh_rent_data <- get_pums(variables = c("TEN",
                                     "FS",
                                     "LNGI",
                                     "GRPIP",
                                     "HINCP",
                                     "NP"),
                       state = "NY",
                       year = 2018,
                       survey = "acs1",
                       recode = TRUE) %>% 
  filter(TEN_label == "Rented") %>% 
  select(TEN_label, FS_label, LNGI_label, GRPIP, HINCP, NP)
```

```{r hh_renter_data display}
hh_rent_data
```

### For Homeowners

Variables:

1. Tenure (TEN), categorical
2. Yearly food stamp/Supplemental Nutrition Assistance Program (SNAP) recipiency (FS), categorical
3. Limited English-speaking household (LNGI), categorical
4. Selected monthly owner costs as a percentage of household income during the past 12 months (OCPIP), categorical in the sense that < 30% is non-burdened and >= 30% is burdened
5. Selected monthly owner costs (SMOCP), continuous
6. Household income (HINCP), continuous
7. Number of persons in this household (NP), continuous

Number of observations: 123,476

```{r hh_own_data, results = 'hide'}
hh_own_data <- get_pums(variables = c("TEN",
                                      "FS",
                                      "LNGI",
                                      "OCPIP",
                                      "HINCP",
                                      "NP"),
                        state = "NY",
                        year = 2018,
                        survey = "acs1",
                        recode = TRUE) %>% 
  filter(!TEN_label == "Rented",
         !TEN_label == "Occupied without payment of rent",
         !TEN_label == "N/A (GQ/vacant)") %>% 
  select(TEN_label, FS_label, LNGI_label, OCPIP, HINCP, NP)
```

```{r hh_own_data display}
hh_own_data
```

### Issues/Questions

The reason I created two datasets is because SMOCP and OCPIP would not apply to renter-occupied households and GRNTP and GRPIP would not apply to owner-occupied households. But perhaps I should eventually combine these two datasets somehow?

[It might be nice to incorporate Employment Status (ESR), Race (RAC1P), and/or perhaps other data that seems to only be available at the person level. Would I need to de-generalize the data / turn every Household observation into _x_ number of Person observations, with _x_ equal to the number of members in the Household? Or match the Household (SERIALNO) with every person (SERIALNO and SPORDER)?

However, following Tuesday's discussion, it may make more sense to stick with household-level analysis and I think I have enough variables to gather interesting conclusions.]

### Saving the Datasets

```{r save as csv}
write_csv(hh_rent_data, "hh_rent_data.csv")
write_csv(hh_own_data, "hh_own_data.csv")
```

