---
title: "Assignment01"
author: "Jessica Tang"
date: "9/6/2020"
output:
  html_document:
   toc: true
   toc_depth: 3
---

## Housing and Food Security in NY State, by Household

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

I used `data("pums_variables")` to see a dataframe of variables available in PUMS.

I would like to see the proportion and types of households in NY State that are burdened and are on food assistance, to determine the prevalence of this double-barrier situation. Although I am primarily interested in renter-occupied households, I would also potentially like to examine this relationship in owner-occupied households. Other variables for type of household are household language and number of members.

For each renter-occupied and owner-occupied household types, the 4 main situations would be:

* not burdened & not on food assistance
* not burdened & on food assistance
* burdened & not on food assistance
* burdened & on food assistance

```{r libraries}
library(tidycensus)
library(tidyverse)
library(dplyr)
```

### For Renters

Sample population: Renter-occupied households in NY State

Variables:

1. Tenure (TEN), categorical
2. Yearly food stamp/Supplemental Nutrition Assistance Program (SNAP) recipiency (FS), categorical
3. Household language (HHL), categorical
4. Gross rent as a percentage of household income past 12 months (GRPIP), continuous
5. Household income (HINCP), continuous
6. Number of persons in this household (NP), continuous

Number of observations: 60,512

```{r hh_renter_data, results = 'hide'}
hh_rent_data <- get_pums(variables = c("TEN",
                                     "FS",
                                     "HHL",
                                     "GRPIP",
                                     "HINCP",
                                     "NP"),
                       state = "NY",
                       year = 2018,
                       survey = "acs1",
                       recode = TRUE) %>% 
  filter(TEN_label == "Rented") %>% 
  select(TEN_label, FS_label, HHL_label, GRPIP, HINCP, NP)
```

```{r hh_renter_data display}
hh_rent_data
```

### For Homeowners

Sample population: Owner-occupied, non-vacant households in NY State

Variables:

1. Tenure (TEN), categorical
2. Yearly food stamp/Supplemental Nutrition Assistance Program (SNAP) recipiency (FS), categorical
3. Household language (HHL), categorical
4. Selected monthly owner costs as a percentage of household income during the past 12 months (OCPIP), continuous
5. Household income (HINCP), continuous
6. Number of persons in this household (NP), continuous

Number of observations: 123,476

```{r hh_own_data, results = 'hide'}
hh_own_data <- get_pums(variables = c("TEN",
                                      "FS",
                                      "HHL",
                                      "OCPIP",
                                      "HINCP",
                                      "NP"),
                        state = "NY",
                        year = 2018,
                        survey = "acs1",
                        recode = TRUE) %>% 
  filter(!TEN_label == "Rented",
         !TEN_label == "Occupied without payment of rent",
         !TEN_label == "N/A (GQ/vacant)") %>% 
  select(TEN_label, FS_label, HHL_label, OCPIP, HINCP, NP)
```

```{r hh_own_data display}
hh_own_data
```

### Issues/Questions

It might be nice to incorporate Employment Status (ESR), Race (RAC1P), and/or perhaps other data that seems to only be available at the person level. Would I need to de-generalize the data / turn every Household observation into _x_ number of Person observations, with _x_ equal to the number of members in the Household? Or match the Household (SERIALNO) with every person (SERIALNO and SPORDER)?

If continuing with analyzing owner-occupied data as well, would I need to exclude owners who own their homes "free and clear"? Sometimes, these households are still burdened (i.e., 30% or more of monthly owner costs as a percentage of household income).

### Saving the Datasets

```{r save as csv}
write_csv(hh_rent_data, "hh_rent_data.csv")
write_csv(hh_own_data, "hh_own_data.csv")
```

